### 2025-04-02 tex/Appendices.tex

% --- Custom Appendix Overview Page ---
% Adds "Appendix" to the Table of Contents
\addcontentsline{toc}{chapter}{Appendix}
% Creates an unnumbered chapter heading for the list
\chapter*{Appendix}
\label{chap:appendix_overview} % Label for reference if needed

% --- List of Manuscripts (Manual Entry) ---
% This section manually lists the manuscripts like in your example.
% You will need to MANUALLY UPDATE THE PAGE RANGES (pp. X - Y)
% after compiling the document fully at least twice.

\noindent % Prevents paragraph indentation
\textbf{Manuscript 1} \hfill \textbf{pp. 59 - 72} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
GALL, C.; OHAN, J.; GLASER, K.; KARSTEN, U.; SCHLOTER, M.; SCHOLTEN, T.; SCHULZ, S.; SEITZ, S. \& KURTH, J. K. (2022): \\
\textit{Biocrusts: Overlooked hotspots of managed soils in mesic environments.} \\
Journal of Plant Nutrition and Soil Science, 185 (6), 745-751. DOI: 10.1002/jpln.202200252.

\vspace{\baselineskip} % Adds a line space between entries

\noindent
\textbf{Manuscript 2} \hfill \textbf{pp. 73 - 111} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
GALL, C.; NEBEL, M.; QUANDT, D.; SCHOLTEN, T. \& SEITZ, S. (2022): \\
\textit{Pioneer biocrust communities prevent soil erosion in temperate forests after disturbances.} \\
Biogeosciences, 19, 3225–3245. DOI: 10.5194/bg-19-3225-2022.

\vspace{\baselineskip}

\noindent
\textbf{Manuscript 3} \hfill \textbf{pp. 112 - 126} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
GALL, C.; NEBEL, M.; SCHOLTEN, T.; THIELEN, S. M. \& SEITZ, S. (in preparation): \\
\textit{On the impact of soil-moss combinations on surface runoff, percolation, soil erosion, and temporal dynamics of soil water content.}

\vspace{\baselineskip}

\noindent
\textbf{Manuscript 4} \hfill \textbf{pp. 127 - 159} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
THIELEN, S. M.; GALL, C.; EBNER, M.; NEBEL, M.; SCHOLTEN, T. \& SEITZ, S. (2021): \\
\textit{Water`s path from moss to soil: A multi-methodological study on water absorption and evaporation of soil moss combinations.} \\
Journal of Hydrology and Hydromechanics, 69 (4), 421-435. DOI: 10.2478/johh-2021-0021.

% --- Include the Manuscript Files as Chapters ---
% The \include command starts each file on a new page and uses the
% chapter numbering defined by \appendix (A, B, C...)
\input{tex/Manuscript-1}
\input{tex/Manuscript-2}
\input{tex/Manuscript-3}
\input{tex/Manuscript-4}
\input{tex/Manuscript-5}

% --- Add Bibliography, Index etc. if needed ---
% \bibliographystyle{plain} % Or your chosen style
% \bibliography{your_bib_file}