### 2025-04-05 tex/Manuscript-3.tex
Update dissertation files with new manuscript details, remove discussion section, and refine formatting

\chapter{Microbial impact on initial soil formation in arid and semiarid environments under simulated climate change}
\label{chap:manuscript3} % Label for potential cross-referencing (start page)