### 2025-04-05 tex/Manuscript-4.tex

\chapter{Microbially induced soil aggregate turnover across different climates and moisture regimes}
\label{chap:manuscript4} % Label for potential cross-referencing (start page)