### 2025-04-05 tex/Manuscript-5.tex

\chapter{Living and decaying roots as regulators of soil aggregation and organic matter formation---from the rhizosphere to the detritusphere}
\label{chap:manuscript5} % Label for potential cross-referencing (start page)