### 2025-04-08 tex/Appendices.tex
Update dissertation structure by adding tables of contents, appendices, and acknowledgements; revise publication entries for accuracy

% --- Appendix Overview Page ---

% Adds "Appendix" to the Table of Contents (if not already done elsewhere)
% Consider if you want this generic entry in addition to specific Manuscript entries.
% If your included manuscripts already add "Appendix A: Title", "Appendix B: Title", etc.
% to the ToC, you might want to comment out the next line.
\addcontentsline{toc}{chapter}{Appendix Overview}

% Creates an unnumbered chapter heading for this overview page
\chapter*{Appendix Overview}
\label{chap:appendix_overview} % Optional label for reference

% Add some vertical space before the list
\vspace{\baselineskip}

% --- List of Manuscripts (Manual Entry) ---
% This section manually lists the manuscripts based on your provided list.
% You MUST MANUALLY UPDATE THE PAGE RANGES (pp. X - Y)
% after compiling the document fully at least twice.

\noindent % Prevents paragraph indentation
\textbf{Manuscript 1} \hfill \textbf{pp. 55 - 84} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
\textbf{Riveras-Muñoz, N.}; Seitz, S.; Witzgall, K.; Rodríguez, V.; Kühn, P.; Mueller, C. W.; Oses, R.; Seguel, O.; Wagner, D.; and Scholten, T. (2022): Biocrust-linked changes in soil aggregate stability along a climatic gradient in the Chilean Coastal Range. \textit{SOIL}, 8, 717--731. DOI: \href{https://doi.org/10.5194/soil-8-717-2022}{10.5194/soil-8-717-2022}.

\vspace{\baselineskip} % Adds a line space between entries

\noindent
\textbf{Manuscript 2} \hfill \textbf{pp. 85 - 116} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
\textbf{Riveras-Muñoz, N.}; Seitz, S.; Gall, C.; Rodríguez, V.; Witzgall, K.; Kühn, P.; Mueller, C.W.; Oses, R.; Seguel, O.; Wagner, D.; Scholten, T. (submitted): Biocrusts as climate-dependent regulators of erosion, water and nutrient cycling. \textit{Geoderma}.

\vspace{\baselineskip}

\noindent
\textbf{Manuscript 3} \hfill \textbf{pp. 117 - 166} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
Rodríguez, V.; Bartholomäus, A.; Witzgall, K.; \textbf{Riveras-Muñoz, N.}; Oses, R.; Liebner, S.; Kallmeyer, J.; Rach, O.; Mueller, C.W.; Seguel, O.; Scholten, T. and Wagner. D. (2024): Microbial impact on initial soil formation in arid and semiarid environments under simulated climate change. \textit{Frontiers in Microbiology}, 15:1319997. DOI: \href{https://doi.org/10.3389/fmicb.2024.1319997}{10.3389/fmicb.2024.1319997}.

\vspace{\baselineskip}

\noindent
\textbf{Manuscript 4} \hfill \textbf{pp. 167 - 208} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
Mitzscherling, J.; Oses, R.; \textbf{Riveras-Muñoz, N.}; Mueller, C. W.; Seguel, O. and Kühn, P.; Scholten, T. and Wagner, D. (2024): Microbially Induced Soil Aggregate Turnover Across Different Climates and Moisture Regimes. SSRN: DOI: \href{https://doi.org/10.2139/ssrn.4740477}{10.2139/ssrn.4740477}.

\vspace{\baselineskip}

\noindent
\textbf{Manuscript 5} \hfill \textbf{pp. 209 - 243} \\ % <-- UPDATE PAGE NUMBERS MANUALLY
Witzgall, K.; Steiner, F.A.; Hesse, B.D.; \textbf{Riveras-Muñoz, N.}; Rodríguez, V.; Teixeira, P.P.C.; Li, M.; Oses, R.; Seguel, O.; Seitz, S.; Wagner, D.; Scholten, T.; Buegger, F.; Angst, G.; Mueller, C.W. (2024): Living and decaying roots as regulators of soil aggregation and organic matter formation---from the rhizosphere to the detritusphere. \textit{Soil Biology and Biochemistry}, 197, 109503. DOI: \href{https://doi.org/10.1016/j.soilbio.2024.109503}{10.1016/j.soilbio.2024.109503}.

% --- End of Manual List ---

% Add a clear page break before the actual manuscripts start, if desired
\cleardoublepage % Ensures next included manuscript starts on a right-hand page